SHARP II: Dark Matter Substructure, Gravitational Lens flux-ratio Anomalies, and the Case of B1555+375: Beware the Effect of Edge-on Disks

### 2015-08-18 papers/B1555_EdgeOn/B1555_letter.tex
rewrite some sentences in B1555 draft

It was first suggested by \citet{Mao1998} that flux-ratio anomalies
observed in radio-loud multiple images of lensed quasars could be
interpreted as a telltale sign of the presence of substructure in lens
galaxies.  Indeed, small perturbations in the gravitational potential
are sufficient to cause flux-ratio anomalies in strong lensed systems
\citep{Bradac02}.  Moreover, unlike in the optical or infrared, in the
radio band, lensed images are neither sensitive to microlensing from
stars, nor to dust extinction. This makes flux-ratio anomalies a promising
tool to detect \textbf{and measure} dark matter substructures \citep{Dalal2002,N13}.  
%At present, the amount of substructure derived from flux-ratio anomalies
%seems to be disagreement with predictions from numerical simulations, 
%with the former requiring a larger fraction of mass in substructure than
%expected \citep{Xu14}. 
\textbf{At present, the amount of substructure derived from flux-ratio anomalies ease the disagremeent with direct-detections and numerical simulations.
Utilizing the surface brightness distribution of extended arcs and Einstein rings
has been introduced by \citet{K05} and \citet{V09}, the fraction of
substructure measured with this \emph{gravitational imaging technique}
is currently in agreement with theoretical expectations \citep{V14a,V12}, although the simulation results show less level of flux anomalies than expected \citep{Xu14}.}
While this could be the result of small sample
statistics, it could also be an indication that flux-ratio anomalies
have a different origin than substructure. \textbf{ISM and microlensing both can produce flux anomalies in stong lens systems, but the effects from ISM show strong frequency dependence and observing in radio band can avoid microlensing from star-size objects \citep{KD04}. These alternative origins are ruled out among known radio loud flux anomlous systems support the detection of substructures. }[OTHER POSSIBLE ORIGONS--SIMONA]
%

@@

As a first trial of modelling, we create a single SIE model to check
the performance of a simple lensing potential. Consistent with
previous studies \citep{Marlow99, Xu14}, a single elliptical
potential model cannot \textbf{reproduce} the flux anomaly in B1555+375.
\textbf{The next step is to examine if an edge-on disk can cause similar perturbations in the strong lensing system  which results in flux ratio anomalies within the merging images. Among most of mass profiles, exponential disk profile best describes the disk component in spiral galaxies because it matches the light distribution of disk \citep{Kee98}. We thus choose a SIE plus exponential disk model (SIE+expdisk) in our study of B1555+375, focusing on the role of its edge-on disk in lens modelling. The free parameters of the models are, SIE: velocity dispersion ($\sigma$), centroid positions, ellipticity ($e$), and position angle ($\theta$), and exponential disk: intrinsic central density ($kappa_0$), centroid positions, ellipticity ($e$), position angle ($\theta$), and scale length ($R_d$). Detail definition of mass profiles in {\tt gravlens} can be found in \citet{Kee01}.}
  We have not included a third component such as an external
shear component or NFW halo into our model because the number of constraints provided by the
image positions and fluxes would not be sufficient to constrain such a
model. However, we note that the image not affected by dust (image C)
in the Keck AO data is consistent with negligible shear (i.e. shear
strength $\Gamma_s=0.002$), when modeled with the pixelized technique
of \citet{V09}.
%We then create two separate models of the lensing potential that both
%incorporate the disk.  In the first we treat the lens as a SIE bulge\textbf{/halo}
%and an exponential edge-on disk (SIE+expdisk). In the second model
%the disk is represented as a very elongated SIE component (2-SIE).

@@

From a theoretical perspective, \citet{Xu14} also find that substructures
cannot explain the observed radio flux anomalies in all the lenses
that were considered. After adding subhaloes from numerical simulations
to these systems, only one out of the eight radio-loud systems
achieves the level of the flux anomalies seen in the observations.  The rest of
them are unlikely (only $1 - 4$ per cent probability) to have \textbf{observed distribution of flux
anomalies} produced only by substructure, including our target B1555+375.
In fact, among those eight radio lenses, B1555+375 is one of the most
anomalous systems, but it is also one of the most challenging systems
to explain by substructures. \citet{Xu14} have concluded that the failure in
predicting flux-ratios is most likely due to oversimplified/improper lens
modelling. They also point out that the SIE component and the external
shear in their model are orthogonal, indicating a possible missing
ingredient in the lens model.
\textbf{Indeed, it's hard to find physical interpretation for an large [IS $\gamma = 0.14$ A LARGE SHEAR?] and orthogonal external shear.
The SHARP $K^{\prime}$ band observations provide another explaination to their result. The edge-on disk crosses the
radio merging double is consistent of this missing component and fully accounts for the observed flux-ratio anomaly.} 
In general, it is well known  \citep[e.g,][]{Ka91} that lens models based only on the locations of lensed
quasar/AGN images have large degeneracies. Therefore, before
concluding that the flux-ratios in any given lens system are a result
of substructure, it is critical to explore a range of smooth models
that are informed by additional observations or to do a comprehensive
analysis such as that conducted by \citet{Xu14}. Otherwise, the derived
substructure fractions should be considered as upper limits.

>>>>

It was first suggested by \citet{Mao1998} that flux-ratio anomalies
observed in radio-loud multiple images of lensed quasars could be
interpreted as a telltale sign of the presence of substructure in lens
galaxies.  Indeed, small perturbations in the gravitational potential
are sufficient to cause flux-ratio anomalies in strong lensed systems
\citep{Bradac02}.  Moreover, unlike in the optical or infrared, in the
radio band, lensed images are neither sensitive to microlensing from
stars, nor to dust extinction. This makes flux-ratio anomalies a promising
tool to detect \textbf{and measure} dark matter substructures \citep{Dalal2002,N13}.  
%At present, the amount of substructure derived from flux-ratio anomalies
%seems to be disagreement with predictions from numerical simulations, 
%with the former requiring a larger fraction of mass in substructure than
%expected \citep{Xu14}. 
\textbf{At present, the amount of substructures derived from flux-ratio anomalies eases the disagremeent with direct detections and numerical simulations.
\emph{Gravitational imaging technique}, by utilizing the surface brightness distribution of extended arcs and Einstein rings,
has been introduced by \citet{K05} and \citet{V09} and gives the measurements of the fraction of
substructure. The result
is currently in agreement with theoretical expectations \citep{V14a,V12} although the simulation results show less level of flux anomalies than expected \citep{Xu14}.}
While this could be the result of small sample
statistics, it could also be an indication that flux-ratio anomalies
have a different origin than substructure. \textbf{ISM and microlensing both can produce flux anomalies in strong lens systems, but not like substurcture pertubations, the effects from ISM show strong frequency dependence and  microlensing from star-size objects is negligible in radio band. These alternative origins have been ruled out among known radio loud flux anomlous systems which implies detection of substructures \citep{KD04}. }[OTHER POSSIBLE ORIGONS--SIMONA]
%

@@

As a first trial of modelling, we create a single SIE model to check
the performance of a simple lensing potential. Consistent with
previous studies \citep{Marlow99, Xu14}, a single elliptical
potential model cannot \textbf{reproduce} the flux anomaly in B1555+375.
\textbf{The next step is to examine if an edge-on disk can cause similar perturbations in the strong lensing system  which results in flux ratio anomalies within the merging images. Among most of mass profiles, exponential disk profile best describes the disk component in spiral galaxies because it matches the light distribution of disk \citep{Kee98}. We thus choose a SIE plus exponential disk model (SIE+expdisk) in our study of B1555+375, focusing on the role of its edge-on disk in lens modelling. The free parameters of the model are, SIE: velocity dispersion ($\sigma$), centroid positions, ellipticity ($e$), and position angle ($\theta$); exponential disk: intrinsic central density ($\kappa_0$), centroid positions, ellipticity ($e$), position angle ($\theta$), and scale length ($R_d$). Detail definition of mass profiles in {\tt gravlens} can be found in \citet{Kee01}.}
  We have not included a third component such as an external
shear component or NFW halo into our model because the number of constraints provided by the
image positions and fluxes would not be sufficient to constrain such a
model. However, we note that the image not affected by dust (image C)
in the Keck AO data is consistent with negligible shear (i.e. shear
strength $\Gamma_s=0.002$), when modeled with the pixelized technique
of \citet{V09}.
%We then create two separate models of the lensing potential that both
%incorporate the disk.  In the first we treat the lens as a SIE bulge\textbf{/halo}
%and an exponential edge-on disk (SIE+expdisk). In the second model
%the disk is represented as a very elongated SIE component (2-SIE).

@@

From a theoretical perspective, \citet{Xu14} also find that substructures
cannot explain the observed radio flux anomalies in all the lenses
that were considered. After adding subhaloes from numerical simulations
to these systems, only one out of the eight radio-loud systems
achieves the level of the flux anomalies seen in the observations.  The rest of
them are unlikely (only $1 - 4$ per cent probability) to have \textbf{observed distribution of flux
anomalies} produced only by substructure, including our target B1555+375.
In fact, among those eight radio lenses, B1555+375 is one of the most
anomalous systems, but it is also one of the most challenging systems
to explain by substructures. \citet{Xu14} have concluded that the failure in
predicting flux-ratios is most likely due to oversimplified/improper lens
modelling. They also point out that the SIE component and the external
shear in their model are orthogonal, indicating a possible missing
ingredient in the lens model.
\textbf{Indeed, it's hard to find physical interpretation for an large and orthogonal external shear.
The SHARP $K^{\prime}$ band observations provide another explaination to their result. The edge-on disk crosses the
radio merging double is consistent of this missing component and fully accounts for the observed flux-ratio anomaly.} 
In general, it is well known  \citep[e.g.][]{Ka91} that lens models based only on the locations of lensed
quasar/AGN images have large degeneracies. Therefore, before
concluding that the flux-ratios in any given lens system are a result
of substructure, it is critical to explore a range of smooth models
that are informed by additional observations or to do a comprehensive
analysis such as that conducted by \citet{Xu14}. Otherwise, the derived
substructure fractions should be considered as upper limits.
